\Large \bf Simple High-Level Code For Cryptographic Arithmetic: \\ With Proofs, Without Compromises

### 2021-08-15 papers/source/rewriting/pldi-2020-reviews-response-rejection.txt

PLDI 2020 Paper #226 Reviews and Comments
===========================================================================
Paper #226 A Framework for Building Verified Partial Evaluators

Review #226A
===========================================================================

Overall merit
-------------
B. I support accepting this paper but will not champion it.

Reviewer expertise
------------------
Z. Some familiarity

Paper summary
-------------
This paper presents a new approach for partial evaluation in Coq. The
goal is to generate more efficient extracted code using partial
evaluation, but most existing partial evaluators for verified code
either increase the size of the trusted code base or cannot scale to
large programs. The new approach combines a different term encoding with
verified (potentially conditional) optimizations. The new framework is
capable of extracted efficient implementations of larger elliptic curve
crypto primitives than previous work.

Comments for author
-------------------
This is a well-written paper that was relatively easy to follow for me,
a total non-expert in interactive theorem proving. It walks the reader
through the motivation for a better partial evaluator (Fiat Crypto
couldn't extract code for larger curves), the desiderata, and the
solutions for each problem that arises. It's hard to argue with the
results -- the new extraction process is much faster, allowing it to
generate code for curves that the old implementation couldn't.

As a non-expert, one challenge I had was that this paper reads more like
a Coq experience report than a new conceptual contribution. There were a
number of places in the paper where I (and, it sounds like, the authors)
wondered if all this effort was just to work around some limitations of
the Coq implementation. Concretely, there are several points (like the
end of Section 3) where it seems like things would have been much easier
if not for termination checking; then, later on, it's revealed that we
got lucky(?) and the chosen PHOAS representation satisfies the
termination checker without any effort. I wonder how much those in the
PLDI audience who don't hack a lot of Coq will be able to learn from
this paper.

Section 3.1 discusses the need for efficient pattern matching for a
partial evaluator, and how pattern matching in Coq's logic isn't very
sophisticated, so the authors rolled their own based on [20]. That makes
sense. However, Section 5.2 concedes that actually, the performance of
the partial evaluator running inside Coq is pretty poor. To get that
performance back up, we need to extract the evaluator to OCaml and run
it natively. It seems like if we're going to make that concession, we
don't need to worry much about the efficiency of the pattern matching
engine -- we can just delegate to OCaml's, just like [1]?

Something I was unclear about throughout the paper was what the user
experience is for a tool like this. One of the cool things about LMS was
that it was trivial to adopt -- just changing the types of the code you
wanted to stage was enough to get the partial evaluation effect. I'm
less clear on the story here -- if I've written and proven an
implementation of some code, and now want to specialize it to a
particular input, what exactly do I have to write and/or do? Section 4.4
left me wondering how often I would actually need to do a lot more than
just write down some rewrite rules -- it seems like depending on the
problem, I would need to (1) write a new abstract interpretation, (2)
prove that correct, and (3) write new rules with the results of that
interpretation in mind. How laborious is that effort?

Post author-response comments
-----------------------------
Thanks for a well-written and helpful response. Ultimately, as a non-expert, I needed more help understanding how this paper would appeal to the PLDI audience broadly. In the paper's current form, I had a hard time distinguishing conceptual insights from Coq engineering tricks.

Review #226B
===========================================================================

Overall merit
-------------
A. I will champion accepting this paper.

Reviewer expertise
------------------
Y. Knowledgeable

Paper summary
-------------
The paper addresses the problem of performing rewrite-intensive program transformation in the context of a provably semantics-preserving system, with proofs generated in checked with a proof assistant. Since type theory-based proof assistants, by design, have limited kernels, and the primitive operations therein are small (e.g. beta-reduction and unfolding of definitions), the proofs quickly become too large too manage and generating them infeasible. The approach combines several techniques to avoid expensive rewriting under binders (such as Coq’s setoid_rewrite) but still use the native kernel for lambda calculus normalization.

Comments for author
-------------------
This work takes a very practical standpoint: if verified toolchains and proof-carrying code are to become mainstream, then someone has to do something about those proof term sizes. The solution is very creative, too. In some sense, it’s even counterintuitive: implement another layer of embedding when what you’re striving for is performance? — but the paper eventually delivers and the way that it does that is clever, by capitalizing on existing normalization mechanics that are already built into Coq.

I find the treatment of the problem systematic and one that presents several different facets of proof-assistant engineering. Also, the writing is superb.

= Specific comments =

p.1 [75] “fancier reduction strategies grow the trusted code base” — this is a bit over-generalizing. They don’t necessarily do so, and I understand after reading this paper that it’s either that or really blowing up the proof size, but IMHO it should be stated as a tradeoff.

p.2 [213] “reduction included in .. need not be requested explicitly.” What does it mean for a reduction to be “requested”?

p.5 [457] “rely on no such embeddings” — it’s perhaps besides the point, but Coq’s extraction mechanism must rely on the semantics of OCaml, and probably any attempt to build a certified extraction procedure will need to have such embedding, right?

p.5 [498] “set of inductive codes” — what are these codes? Is it just an implementation detail?

p.5 [545] “does not need to look into definitions” — so this rewriting scheme never unfolds the definitions of symbols? Why is that an acceptable limitation?

p.6 [630] “Wildcard”. The holes are not named nor indexed. Is it irrelevant or are they implicitly indexed left-to-right?
p.6 [636] “App” — I am guessing the edge in the decision tree should read “Switch” (as in [609])? Or is “Switch” supposed to be written inside the circle?

p.6 [654] “match e with etc.” — Where did the vectors go in this expression? Were they all eliminated by Coq’s Compute or similar?

p.11 [1169] “including reification, cbv, etc.” — It is not clear how important these are, and whether the reader should consider them as part of the running time or not. Surely reification is a vital part of the process, so what is the significance of “only rewriting”?

p.11 [1205] “Macrobenchmark”: maybe it’s just me but I read this as “Microbenchmark” (following 5.1) and was confused.

Post author-response comments
-----------------------------
*Outcome of PC discussion*

While all the reviewers appreciated the effort involved in leveraging the agreeably weak beta-reduction in Coq’s kernel without resorting to OCaml implementation, the contribution is found to be too narrow and the vast majority of the PLDI audience will not be able to relate to it.

Review #226C
===========================================================================

Overall merit
-------------
B. I support accepting this paper but will not champion it.

Reviewer expertise
------------------
Z. Some familiarity

Paper summary
-------------
Partial evaluation is an optimization technique that aims at generating faster, specialized variants of a function by reducing its application domain. An example in cryptography is a specialization of a function to a certain key size. The resulting function is faster than the more generic function, but a proof is required to demonstrate that the behaviour of the new function is the same as the general function. This paper presents a way to efficiently prove the correctness of partial evaluators. Although previous approaches exist, they introduced a trusted component in the Trusted Computing Base (TCB), and had performance issue due to the lack of common sub-expression handling. This paper proposes an approach that does not add to the TCB and that takes common sub-expression into account to reduce the redundancy in generated proofs.

In order to prove correctness of partial evaluators, the authors have developed a Coq library, which works in two steps: first a rewriter is generated from equality lemmas, and then it can be applied to prove the equality of two functions or synthesize a specialized version of a function with its correctness proof. The paper presents a quick overview of the features of the library, followed by a more in-depth explanation of the techniques used to build it, comparing with existing implementations, as their work is heavily inspired by work by Aehlig et al. The paper evaluates the library speed compared to coq standard tactics on two microbenchmarks, and on a macrobenchmark involving code generation for the Fiat cryptography library.

Comments for author
-------------------
Positives
----------- 
- The paper is well-structured and easy to follow. 
- The authors are positioning well their work in the context of existing approaches and techniques that they use. The paper explains when they adapted them and justifies all of their choices. 
They also compared how their own use-case impacts existing approaches, and how their implementation overcomes the difficulty. Overall, the formalization flows almost naturally.

Negatives
---------------

- Through the paper there is a constant comparison between the introduced  approach and the work of Aehlig et al [1]. One expects that those two approaches will also be empirically compared and evaluated. 
- Although important to the presented work, side conditions (in Sections 4.3 and 4.4) are not as well explained and justified as the rest of the paper. 
- The abstract interpreter is not evaluated.

Suggestions
----------------------------------------------
- Figure 3 (benchmark results) is really hard to read and evaluate since the text is small and the graphs are barely readable. Slightly reducing the size of Figure 1 might help you get enough space for Figure 3.

Review #226D
===========================================================================

Overall merit
-------------
B. I support accepting this paper but will not champion it.

Reviewer expertise
------------------
Y. Knowledgeable

Paper summary
-------------
This paper proposes a new approach to do partial evaluation in Coq.  A large component of it is a term rewriter, which is used to simplify terms as knowledge about some parameters is given.  Unlike previous work by Achlig etal [1], where an operational semantics of an ML-like language for terms was defined via a deep embedding and had to be trusted as part of the overall proof checking, here the authors propose an approach where the language is defined at the core level via translation back and forth to Coq itself.  While this translation-based definition still needs to be given, the advantage is that Coq can now take over a lot of the checking work (like termination) and thus the overall trust base stays the same, without any need to modify or extend the proof checker.

Coq being Coq, a lot of challenges had to be overcome in order to achieve usable performance.  First, a less-than-naive rewrite engine had to be implemented, where some limited indexing across rewrite rules was used, to avoid doing work from scratch each time a rule is attempted to be matched.    Second, translations to Coq and normalization of evaluation of terms had to be implemented in a certain way, to convince Coq that everything terminates.  Third, a shallow encoding of terms' binders into Coq's existing binders had to be implemented, to avoid the cost of de Bruijn encodings.  Then subterm sharing and side conditions to rules were necessary, which are also considered basic features of and implemented in other rewrite engines outhere.

Evaluation was done both on microbenchmarks and on a larger example, the Fiat Crytocurrency.  The results show good results.

Comments for author
-------------------
The paper is overall well-written.  The choice of what to include in the paper and what in the appendix is fair, although there is a lot of material in the appendix that is required to appreciate the contribution.  Not sure what you can do about it, though, but at least make sure that it will be all available in a companion report if the paper is accepted.  I do not have any major questions, only a few minor ones that made me have some reservations about the paper and not give it an A.

Why not implement, once and for all in Coq, the efficient rewriting engine algorithms, based on indexing and pattern match automata, that the term rewriting community has developed over the years and implemented in truly fast rewrite engines, performing millions, or even tens of millions, or rewrite steps per second?  That would benefit the entire Coq community.

Why should one trust the rewrite rules as correct?  Don't they become part of the trust base anyway?  Also, isn't your term translation back and forth to Coq the same as like giving a semantics to an ML-like language like in [1]?  One still has to trust these translations, including the fact that the ML-like language and Coq share the semantics of their similar constructs., which to me counts as the same level of trust that [1] requires from their users.  You should clarify that in the paper.

Lines 701-704: this becomes part of the trust base, no?

How is the approach in [11] different from the approach here?  Is it only a performance improvement?  How about the trust-base?  I was a bit disappointed to read that you generate essentially the same code as [11] and you only care about the performance of the partial evaluator; it would have been a really nice application where correctness is guaranteed through Coq, but if the application already had all these guarantees, then the work in this paper feels less critical and more incremental.

Response by Jason Gross <jgross@mit.edu>
---------------------------------------------------------------------------
Thank you all for your feedback and comments.  We'll start by
clarifying a few most crucial elements of our approach and its
advantages, after which we'll return to answer further questions on a
review-by-review basis (some of which are forward-referenced by our
initial highlights).

First, a quick response to Review A's questions about what the user
experience is really like.  Section 1.1 does provide a complete
example; we are not eliding any additional user overhead of applying
our rewrite engine.  (Abstract interpretation is a separate activity
that we say a bit more about later, but this paper is not about
abstract interpretation, but rather a flexible framework that can
interface with abstract intepretation.)

Next: Review D revealed a number of places where we left it unclear
how we avoid growing the trusted base, beyond what one accepts anyway
in using Coq, and we're glad to have the opportunity to clarify.
Briefly:
- *Must one trust that the registered rewrite rules are correct*?
   No, the verified rewriting engine requires a proof that the
   interpretation of the AST of each rewrite rule is a correct
   proposition in the logic.
- *Must one trust the translations back and forth between normal Coq and our ASTs?*
  No, the translation *into* ASTs is inherently untrusted, as we ask
  the Coq kernel to "run it backwards" through interpretation of ASTs,
  for instance in formulating the semantic correctness condition for a
  rewrite rule. The kernel naturally checks that interpretation takes
  us back to the terms (e.g., rewrite-rule statements) we started
  with.
- *Are the central NbE function definitions (lines 701-704) trusted?*
  No, they receive full proofs of functional correctness (which are
  used as lemmas in the overall engine proof) and thus need not be
  trusted.
Summing up, in any particular use of our rewrite engine, the final
theorem statement proved truly does not mention or depend on any part
of our framework; only the *proof* depends on our framework.  Since we
do not add any axioms in this project, the trusted base is not
expanded.  And see our more detailed Review D response below for more.

Finally, a few brief remarks on performance, which indeed is at the
heart of our project's motivation. Review D expresses disappointment
that, in comparison to [11], our project is only about improving the
performance of a compilation pipeline; we should have emphasized that
the tool from [11] timed out (running longer than 100 hours) on many
important inputs, while our new version handles all known real-world
elliptic curves usually within a few minutes or at most two hours. Also,
Review A summarizes Section 5.2 as concluding that our engine's
performance is poor. We want to emphasize again that there we compare
against reduction on programs (from [11]) that have been
tailor-written in CPS just to improve reduction performance, while
our engine works on much more readable, direct-style code, in addition
to integrating flexible rewriting with reduction.  Still, our
within-Coq evaluation doesn't pass 90 seconds of running time until
primes grow to about 10 times as many bits as native machine words.

Responses to Review A
=====================

Regarding termination checking, indeed, normalization by evaluation in
PHOAS is, in some sense, specifically crafted to appease the
termination checker in an elegant way.

Broadly, none of the individual techniques are new.  What's new is
that we combine all of them to get a flexible verified tool that
operates efficiently at scale, without extending the trusted code
base.

> However, Section 5.2 concedes that actually, the performance of the
> partial evaluator running inside Coq is pretty poor.

While the performance is poor compared to the extracted version and is
not stellar compared to the "rewrite all the code in CPS" version
(which is inflexible), the performance is still quite good relative to
the prior state-of-the-art equational reasoning in Coq of
`setoid_rewrite` / `rewrite_strat`, which take around 90 seconds for
primes equal to the bitwidth, and about 3 hours for primes with twice
as many bits as the machine architecture.  By contrast, even within
Coq, our approach only approaches 90 seconds for primes with about 10
times as many bits as the machine architecture.  We'll aim to
emphasize this more in future versions of the paper.

> It seems like if we're going to make that concession, we don't need
> to worry much about the efficiency of the pattern matching engine --
> we can just delegate to OCaml's, just like [1]?

If we delegate to OCaml's pattern matching engine, then we can't do
conditional rewrite rules, because there's not currently a way to
extract Coq pattern matches to OCaml's conditional matches.
Additionally, we tried eliminating pattern-matching compilation from
our engine.  This results in out-of-memory errors during compile-time
(because we end up being quadratic in the number of rewrite rules for
reasons that are slightly hard to explain), and if we disable the
partial evaluation of rewrite rule selection (trusting in OCaml's
pattern matching engine), we lose approximately a factor of 8 in
performance of extracted code.

The intended user experience for a tool like this is detailed at the
end of section 1.1.  The couple lines of code described there are all
that is necessary for equational reasoning.  The work of abstract
interpretation was required for Fiat Cryptography to handle automatic
specialization from unbounded integers to fixed-width integers.  This
is more than just equational reasoning, and, indeed, was a somewhat
laborious process.  We have yet to find any better alternative to
this, however, when abstract interpretation is, in fact, required.

Responses to Review B
=====================

Thank you for your detailed comments about places where we were
unclear or misstating things.

Responding to the questions:

> p.2 [213] “reduction included in .. need not be requested
> explicitly.” What does it mean for a reduction to be “requested”?

Imagine a proof assistant without any computational reduction rules,
only axioms.  In such a system, every reduction step needs to
correspond to a use of an axiom in the proof term.  This is what we
mean by "requested".

> p.5 [457] “rely on no such embeddings” — it’s perhaps besides the
> point, but Coq’s extraction mechanism must rely on the semantics of
> OCaml, and probably any attempt to build a certified extraction
> procedure will need to have such embedding, right?

Yes.

> p.5 [498] “set of inductive codes” — what are these codes? Is it
> just an implementation detail?

The inductive codes are the inductive types from step (2), "Inductive
types enumerating all available primitive types and functions are
emitted."  For example, we might have the inductive codes for types
`Inductive base := nat_code | bool_code | Z_code` and the inductive
codes for functions `Inductive ident := nat_add | Z_add | bool_if`.
Details beyond this are basically just implementation details.

> p.5 [545] “does not need to look into definitions” — so this
> rewriting scheme never unfolds the definitions of symbols? Why is
> that an acceptable limitation?

The rewriter itself has no native support for unfolding definitions,
but you can include a rewrite rule that corresponds to an "unfolding",
such as `?n + S ?m = S (n + m)` or `fst (?x, ?y) = x`.

> p.6 [630] “Wildcard”. The holes are not named nor indexed. Is it
> irrelevant or are they implicitly indexed left-to-right?

Holes are indexed left-to-right, but this is an implementation detail;
the machinery for reifying Gallina rewrite rules into patterns and
replacements takes care of indexing.

> p.6 [636] “App” — I am guessing the edge in the decision tree should
> read “Switch” (as in [609])? Or is “Switch” supposed to be written
> inside the circle?

We should have been more clear that every un-labeled node is a
"Switch" node.  The "App" edges are for the `app_case` on [609], and
the identifier-labeled edges are part of `icases`.

> p.6 [654] “match e with etc.” — Where did the vectors go in this
> expression? Were they all eliminated by Coq’s Compute or similar?

Yes, they are all eliminated under `cbv` / `cbn` during the partial
evaluation described on lines [644]-[646].

> p.11 [1169] “including reification, cbv, etc.” — It is not clear how
> important these are, and whether the reader should consider them as
> part of the running time or not. Surely reification is a vital part
> of the process, so what is the significance of “only rewriting”?

### 2021-08-15 _site/papers/source/fiat-crypto/sigops-osr-2020/fiat-crypto.tex
Add back deploy

\documentclass[conference,letterpaper]{IEEEtran}
%\overfullrule=1mm
\bibliographystyle{IEEEtran}

@@

% packages for the carry-chain diagram
\usepackage{wrapfig}
\usepackage[all, frame]{xy}
\usepackage{mathtools}
\include{digitboxed}

@@

\begin{document}

\title{Simple High-Level Code For Cryptographic Arithmetic -- With Proofs, Without Compromises}

\author{%
 \IEEEauthorblockN{Andres Erbsen \quad Jade Philipoom \quad Jason Gross \quad Robert Sloan \quad Adam Chlipala}
 \IEEEauthorblockA{%
   MIT CSAIL, \\
   Cambridge, MA, USA \\
   \{\texttt{andreser}, \texttt{jadep}, \texttt{jgross}\}\texttt{@mit.edu}, \texttt{rob.sloan@alum.mit.edu}, \texttt{adamc@csail.mit.edu} %TODO(@csail.mit.edu for all?)
 }
}

@@

\begin{abstract}
  We introduce a new approach for implementing cryptographic arithmetic in short high-level code with machine-checked proofs of functional correctness.
  We further demonstrate that simple partial evaluation is sufficient to transform such initial code into the fastest-known C code,
  breaking the decades-old pattern that the only fast implementations are those whose instruction-level steps were written out by hand.

  These techniques were used to build an elliptic-curve library that achieves competitive performance for 80 prime fields and multiple CPU architectures, showing that implementation and proof effort scales with the number and complexity of conceptually different algorithms, not their use cases.
  As one outcome, we present the first verified high-performance implementation of P-256, the most widely used elliptic curve.
  Implementations from our library were included in BoringSSL to replace existing specialized code, for inclusion in several large deployments for Chrome, Android, and CloudFlare.
\end{abstract}

\IEEEpeerreviewmaketitle

\section{Motivation}

The predominant practice today is to implement cryptographic primitives at the lowest possible level of abstraction: usually flat C code, ideally assembly language if engineering time is plentiful.
Low-level implementation has been the only way to achieve good performance, ensure that execution time does not leak secret information, and enable the code to be called from any programming language.
Yet placing the burden of every detail on the programmer also leaves a lot to be desired.
The demand for implementation experts' time is high, resulting in an ever-growing backlog of algorithms to be reimplemented in specialized code ``as soon as possible.''
Worse, bugs are also common enough that it is often not economical to go beyond fixing the behavior for problematic inputs and understand the root cause and impact of the defect.

There is hope that these issues can eventually be overcome by redoubling community efforts to produce high-quality crypto code.
However, even the most renowned implementors are not infallible, and if a typo among thousands of lines of monotonous code can only be recognized as incorrect by top experts,\footnote{We refer to the \texttt{ed25519-amd64-64-24k} bug that is discussed along with other instructive real-world implementation defects in Appendix \ref{bugsapp}.} it is reasonable to expect diminishing returns for achieving correctness through manual review.
A number of tools for computer-aided verification have been proposed~\cite{verif25519,gfverif,ECCstar,hacl,vale,jasmin}, enabling developers to eliminate incorrect-output bugs conclusively.
Three solid implementations of the X25519 Diffie-Hellman function were verified.
In the beautiful world where X25519 was the only arithmetic-based crypto primitive we need, now would be the time to declare victory and go home.
Yet most of the Internet still uses P-256, and the current proposals for post-quantum cryptosystems are far from Curve25519's combination of performance and simplicity.

A closer look into the development and verification process (with an eye towards replicating it for another function) reveals a discouraging amount of duplicated effort.
First, expert time is spent on mechanically writing out the instruction-level steps required to perform the desired arithmetic, drawing from experience and folklore.
Then, expert time is spent to optimize the code by applying a carefully chosen combination of simple transformations.
Finally, expert time is spent on annotating the now-munged code with assertions relating values of run-time variables to specification variables, until the SMT solver (used as a black box) is convinced that each assertion is implied by the previous ones.
As the number of steps required to multiply two numbers is quadratic in the size of those numbers, effort grows accordingly: for example ``Verifying Curve25519 Software''~\cite{verif25519} reports that multiplication modulo $2^{255}-19$ required 27 times the number of assertions used for $2^{127}-1$.
It is important to note that both examples use \emph{the same} simple modular multiplication algorithm, and the same correctness argument should apply, just with different parameters.
This is just one example of how manual repetition of algorithmic steps inside primitive cryptographic arithmetic implementations makes them notoriously difficult to write, review, or prove correct.

Hence we suggest a rather different way of implementing cryptographic primitives:
to implement general algorithms in the style that best illustrates their correctness arguments
and then derive parameter-specific low-level code through partial evaluation.
For example, the algorithm used for multiplication modulo $2^{127}-1$ and $2^{255}-19$ is first encoded as a purely functional program operating on lists of digit-weight pairs: it generates a list of partial products, shifts down the terms with weight greater than the modulus, and scales them accordingly.
It is then instantiated with lists of concrete weights and length, for example ten limbs with weights $2^{\lceil 25.5 i\rceil}$, and partially evaluated so that no run-time manipulation of lists and weights remains, resulting in straight-line code very similar to what an expert would have written.
As the optimization pipeline is proven correct once and for all, correctness of all code generated from a high-level algorithm can be proven directly on its high-level implementation.
A nice side benefit of this approach is that the same high-level algorithm and proof can be used for multiple target configurations with very little effort, enabling the use of fast implementations in cases that were previously not worth the effort.
Finally, the generated code is compiled with a conventional C compiler.
As current compiler support for different CPU features varies wildly in optimization quality, this step can limit the performance of our code (especially when the best compilation would make careful use of carry flags or vector instructions).
We aim for the fastest code in terms of ALU functions, but we leave instruction selection, scheduling, and register allocation out of scope.

In practice, we use the fast arithmetic implementations reported on here in conjunction with a library of verified cryptographic algorithms, the rest of which deviates much less from common implementation practice.
All formal reasoning is done in the Coq proof assistant, and the overall trusted computing base also includes a simple pretty-printer and the C language toolchain.
The development, called \textbf{Fiat Cryptography}, is available under the MIT license at:
\begin{center}
  \texttt{\url{https://github.com/mit-plv/fiat-crypto}}
\end{center}

The approach we advocate for is already used in widely deployed code.
Most applications relying on the implementations described in this paper are using them through BoringSSL, Google's OpenSSL-derived crypto library that includes our Curve25519 and P-256 implementations.
A high-profile user is the Chrome Web browser, so today about half of HTTPS connections opened by Web browsers worldwide use our fast verified code (Chrome versions since 65 have about 60\% market share~\cite{browserMarketShare}, and 90\% of connections use X25519 or P-256~\cite{davidenTLSECC}).
BoringSSL is also used in Android and on a majority of Google's servers, so a Google search from Chrome would likely be relying on our code on both sides of the connection.
Maintainers appreciate the reduced need to worry about bugs.

>>>>

%Page: single-spaced 8.5″ x 11″ pages
%Font and Columns: 10-point type on 12-point (single-spaced) leading, two-column format, Times Roman or a similar font, within a text block 6.5″ wide x 9″ deep.
%Number of pages: OSR does not enforce strict minimum or maximum page limits. We suggest 2-6 pages as a rough guideline, but authors should feel free to devote more space as needed..
%Other considerations: Pages should NOT be numbered, and figures and tables should be legible in black and white, without requiring magnification.
%Please feel free to use any Systems conference template (e.g., USENIX ATC, OSDI, SOSP, NSDI, etc.) to prepare drafts.

\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix2019_v3}

\bibliographystyle{plain}
\pagestyle{empty}

@@

% packages for the carry-chain diagram
\usepackage{wrapfig}
\usepackage[all, frame]{xy}
\usepackage{mathtools}

@@

%-------------------------------------------------------------------------------
\begin{document}
%-------------------------------------------------------------------------------

%don't want date printed
\date{}

% make title bold and 14 pt font (Latex default is non-bold, 16 pt)
\title{\Large \bf Simple High-Level Code For Cryptographic Arithmetic: \\ With Proofs, Without Compromises}

% The template didn't include email addresses; do we want to?
\author{
{\rm Andres Erbsen} \\
MIT CSAIL \\
{\rm \texttt{andreser@mit.edu}}
\and
{\rm Jade Philipoom} \\
MIT CSAIL \\
{\rm \texttt{jadep@mit.edu}}
\and
{\rm Jason Gross} \\
MIT CSAIL \\
{\rm \texttt{jgross@mit.edu}}
\and
{\rm Robert Sloan} \\
MIT CSAIL \\
{\rm \texttt{rob.sloan@alum.mit.edu}}
\and
{\rm Adam Chlipala} \\
MIT CSAIL \\
{\rm \texttt{adamc@csail.mit.edu}}
} % end author

@@

\begin{abstract}
  We introduce an unusual approach for implementing cryptographic arithmetic in short high-level code with machine-checked proofs of functional correctness.
  We further demonstrate that simple partial evaluation is sufficient to transform such initial code into highly competitive C code, breaking the decades-old pattern that the only fast implementations are those whose instruction-level steps were written out by hand.

These techniques were used to build an elliptic-curve library that achieves competitive performance for a wide range of prime fields and multiple CPU architectures, showing that implementation and proof effort scales with the number and complexity of conceptually different algorithms, not their use cases.
As one outcome, we present the first verified high-performance implementation of P-256, the most widely used elliptic curve.
Implementations from our library were included in BoringSSL to replace existing specialized code, for inclusion in several large deployments for Chrome, Android, and CloudFlare.

This is an abridged version of the full paper originally presented in IEEE S\&P 2019~\cite{FiatCryptoSP19}.
We have omitted most proof-engineering details in favor of a focus on the system's functional capabilities.
\end{abstract}

\section{Introduction}

As verification makes the transition from research topic to component of real-world software ecosystems, the field must face challenges of scale, in terms of both amount of code and conceptual complexity.
Each system, of course, has its own constraints and objectives, and these needs will best be met with a broad toolbox of approaches.
In this paper, we sketch a case study following a relatively unusual approach, in the hope that it will help illuminate the broad array of possibilities for verification projects.
The full paper was originally presented in IEEE S\&P 2019~\cite{FiatCryptoSP19}; what follows is a version abridged, selectively updated, and adapted for a more systems-oriented audience.

That audience is probably most familiar with projects fitting the classic mold of proving theorems about programs written by hand, as exemplified by sEL4~\cite{seL4SOSP09}, IronClad~\cite{hawblitzel2014ironclad,hawblitzel2015ironfleet}, FSCQ~\cite{FscqSOSP15,FscqSOSP17,FscqOSDI18}, CertiKOS~\cite{CertiKOS}, and various case studies in push-button verification~\cite{Pbfs,Hyperkernel,Nickel,Serval}.
Often the implementers of these projects do rewrite systems from scratch in tandem with proofs, but the final code artifacts implement very similar pseudocode to off-the-shelf systems.
A moderately widely known alternative is \emph{program synthesis}, which writes a program automatically from its logical specification.
Most colloquially, this idea is associated with relatively intensive combinatorial search through spaces of possibilities, often taking advantage of SMT solvers like Z3~\cite{Z3}.
However, some systems-code domains are focused enough that we can write very predictable and \emph{verified} generic implementations.
With the help of a verified domain-specific compiler, we can translate these general problem descriptions into practical, performance-competitive code.
This paper considers a successful example of the template-based code-generation strategy.

Our project verifies efficient modular-arithmetic routines for elliptic-curve cryptography, a domain to which much optimization effort has been applied.
This means that competitive implementations use highly specialized techniques, previously accessible only to a small number of experts.
We aimed to create a system that generalized these approaches and proved, with minimal trusted code, that the detailed low-level routines implemented their high-level arithmetic specifications.
To that end, we first studied handwritten implementations and extracted fully general descriptions of the complex design choices they required.
Then, we encoded these algorithms in the Coq proof assistant~\cite{coq} as templates, generic over compile-time parameters like the modulus and integer width of the underlying architecture.
Once these parameters are plugged in, a domain-specific compiler transforms the templates into efficient low-level code.

The output code is encoded in a specialized low-level abstract syntax tree (AST) that includes only very basic fixed-width integer operations, at the approximate abstraction level of assembly instructions, and is easily translatable to a number of output languages.
At current count we can produce C, Go, and Rust, along with two other unusual specialized languages for specific applications.
The C code, our primary target, is competitive with existing handwritten C, although it still falls short of the performance of handwritten assembly.
However, our toolchain, with no input other than the compile-time parameters, produces efficient and completely verified code, and we believe it presents a very interesting case study in a verification approach that is unusual in \emph{generating} code, rather than working backwards from existing code.
(It is worthwhile to note here that one can still verify specific existing code by proving it equivalent to the generated code, and the equivalence is much more straightforward than equivalence with a high-level specification, and we have used this technique to verify existing production code.)
Our verification stretches all the way from very low-level C to the natural high-level specifications of modular arithmetic (e.g. $\forall x, y.\ \text{add}(x, y) = (x + y) \bmod p$).
The output code uses a restricted set of instructions by construction, preventing timing side channels.
The overall trusted computing base includes the Coq proof checker, a simple pretty-printer, and the C language toolchain.

In addition to being interesting research, our approach has proven to be practically applicable, as it is already used in widely deployed code.
Most applications relying on the implementations described in this paper are using them through BoringSSL, Google's OpenSSL-derived crypto library that includes our implementations of Curve25519 and P-256, the two most widely used elliptic curves via Internet standards.
A high-profile user is the Chrome Web browser, so today about half of HTTPS connections opened by Web browsers worldwide use our fast verified code (Chrome has about 60\% market share~\cite{browserMarketShare2020}, and 90\% of connections use X25519 or P-256~\cite{davidenTLSECC}).
% todo
% now Edge is just rebranded chrome, so +9% if we can find a citation
% Safari also uses BoringSSL, so +15%, but again nothing to cite, and Apple also may have put their own crypto implementations into BoringSSL and just kept the TLS parts
BoringSSL is also used in Android and on a majority of Google's servers, so a Google search from Chrome would likely be relying on our code on both sides of the connection.
Maintainers appreciate the reduced need to worry about bugs.

We believe our approach is particularly suited to design domains that a) involve methodical low-level transformations that are nonetheless finicky and easy for humans to get wrong, and b) apply these transformations to a wide range of setups with different parameters.
Elliptic-curve field arithmetic fits this description, because most efficient implementations use the same family of optimization techniques, while there are hundreds of combinations of moduli and native-integer sizes.
A change to either one usually dramatically changes the resulting code, even though the underlying techniques are similar.
Other opportunities to apply these ideas may include families of very closely related data structures being used in OS kernels and other systems software.

This abridged paper will be a whirlwind tour through the most interesting and informative aspects of our approach.
First, Section \ref{templates} will go over a common optimization technique for modular reduction and integer representation, as an example of the kinds of algorithms our templates encode.
Then, in Section \ref{compiler}, we will explain the stages of our domain-specific compiler and its capabilities.
Finally, Section \ref{experiments} will show a sample of our benchmarks.
For more details, particularly for proof-engineering techniques, explorations of related work, and full benchmarks, please look to the full-length paper~\cite{FiatCryptoSP19}.
For even \emph{more} details, the entire development is called \textbf{Fiat Cryptography}, and is available under the MIT license at:
\begin{center}
  \texttt{\url{https://github.com/mit-plv/fiat-crypto}}
\end{center}

\section{Arithmetic Template Library}\label{templates}

In this section, we will describe the arithmetic optimization strategies we encoded in our template library.
For those who prefer to read code, we suggest \texttt{src/Demo.v} in the framework's source code, which contains a succinct standalone development of the unsaturated-arithmetic library up to and including an implementation of modular reduction specialized to take advantage of the specific shape of its fixed modulus.
The concrete examples derived in this section are within established implementation practice, and an expert would be able to reproduce them given an informal description of the strategy.
Our contribution is to encode this general wisdom in concrete algorithms and provide correctness certificates.

Cryptographic modular arithmetic implementations distribute very large numbers across smaller \emph{limbs} of 32- or 64-bit integers.
Fig. \ref{fig:representationexamples} shows a small sample of fast representations for different primes.
Notice that many of these implementations use bases other than $2^{32}$ or $2^{64}$, leaving bits unused in each limb: these are called \emph{unsaturated} representations.
Conversely, the ones using all available bits are called \emph{saturated}.
This abridged paper will only describe the unsaturated arithmetic in our development; for saturated arithmetic, please refer to the full paper.

Why accept all of this added complication, and waste available bits, instead of just using saturated representations for everything?
These unorthodox big integers are fast primarily because of a specific modular-reduction algorithm, pseudo-Mersenne reduction, which becomes extremely fast when the number of bits in the prime \emph{corresponds to a limb boundary}.
For instance, reduction modulo $2^{255}-19$ is fastest when the 255th bit of a bignum is the first bit of a limb.
In a saturated implementation, limb boundaries would fall on multiples of the integer size (for instance, a 64-bit saturated representation would have limb boundaries at bits 0, 64, 128, etc.).
But in an unsaturated implementation, with a funky base like $2^{25.5}$, we get limb boundaries at bits 0, 26, 51, 77, ... and 255.
The vast speed improvement in modular reduction is worth keeping a few extra registers around to represent each large number.

\subsection{A Note on Carrying}

To give a more concrete sense of how the representation described above works in practice, we will walk through a modular multiplication procedure specialized to the (relatively) small modulus $2^{127} - 1$ and a 64-bit architecture.
Say we want to multiply two numbers $s$ and $t$ in its field, with those inputs broken up into limbs as $s = s_0 + 2^{43} s_1 + 2^{85} s_2$ and $t = t_0 + 2^{43} t_1 + 2^{85} t_2$.
Distributing multiplication repeatedly over addition gives us the answer form shown in Fig. \ref{distribute}.

We have formatted the calculation suggestively: down each column, the powers of two are very close together, differing by at most one.
Therefore, it is easy to add down the columns to form our final answer, split conveniently into digits with integral bit widths.
%The number of primitive addition and multiplication operations is reasonable (and the primitive multiplications can be implemented as self-additions, which are often faster in hardware).
